---
title: "Initial Korean covid 19"
author: "Yessica garay"
date: "Springs 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

#Introduction

H$_O There is no difference in death rate based on sex.


H$_A There is a difference in death rate based on sex.

 
#Methods

variables for this study are sex in result

We want to know if the result depends on sex result is a categorical variable.
Sex is also a categorical variable

So we use Cat~Cat helper file

#Results

## Graphical results - use ggplot commands to do this  


```{r}
barchartGC(~result + sex ,data=kRdata)
barchartGC(~result + sex ,data=kRdata,
type="percent")

```

The chart show the mens dies more than womens having the 60% more than womens with the 40%


## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
 
The table shows the average males deaths are higher than female deaths 64% versus females at 36.

## Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```






## Conclusions

