---
title: "Initial Korean covid 19"
author: "Yessica garay"
date: "Springs 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

#Introduction

H$_O There is no difference in death rate based on sex.


H$_A There is a difference in death rate based on sex.

 
#Methods

variables for this study are sex in result

We want to know if the result depends on sex result is a categorical variable.
Sex is also a categorical variable

So we use Cat~Cat helper file

#Results

## Graphical results - use ggplot commands to do this  


```{r}
barchartGC(~result + sex ,data=kRdata)
barchartGC(~result + sex ,data=kRdata,
type="percent")

```

The chart show the mens dies more than womens having the 60% more than womens with the 40%


## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
 
The table shows the average males deaths are higher than female deaths 64% versus females at 36.

## Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi-squared test p-value is 0.0012 wich is lesser than 0.05 which means we can effective reject the null hypothesis so there is a difference between death rate based on sex.
Overall, the results indicate a statistically significant association between the variables, with an estimated odds ratio of 2.3 and a 95% confidence interval ranging from 1.4 to 3.9.


## Conclusions

The original questions is if there is a difference in korea data death based on sex.  The low p-values obtained from both tests (0.001 for Pearson's Chi-squared test and 0.0009 for Fisher's Exact Test) indicate that the association is statistically significant.
Therefore, we reject the null hypothesis of no association between sex and result. Instead, we conclude that there is a significant association between the two variables, with females and males showing different outcomes. 

